A partisan affair: Mapping patronage in municipal bureaucracies of Brazil

### 2021-03-25 paper/draft.tex
include additional info on proportion of party members

\begin{figure}[H]
    \centering
    \includegraphics[width = 8cm, height = 8cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. Sample includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}

@@

Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors. These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatut'{a}rio}). Moving down the hierarchy, it is clear that party members predominantly occupy high level positions, with a much larger prevalence of party members in bureaucrat low, frontline high and then frontline low, as compared to other positions in the bureaucracy.

\begin{figure}[H]
    \centering
    \includegraphics[width = 8cm, height = 8cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
    \caption{\textbf{Proportion of party-affiliated members by occupation}. Darker colors indicate a larger degree of partisan affiliation. Sample includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:partisan_by_cbo}
\end{figure}

>>>>

\begin{figure}[H]
    \centering
    \includegraphics[width = 9cm, height = 9cm]{figures/maps/pooled.png}
    \caption{\textbf{Proportion of party-affiliated members by municipality (2015)}. Darker colors indicate a larger degree of partisan affiliation. Sample includes all bureaucrats, from high-level managers to service staff.}
    \label{fig:map_pooled}
\end{figure}

@@

Party members are relatively concentrated in high level bureaucrat positions, such as managers and directors, as illustrated by figure \ref{fig:partisan_breakdown}. While there is wide variation with respect to the proportion of partisans, the median proportion of party members in top-level bureaucrat positions is 28 percent in the time period between  These are also the best paying positions and the ones with the highest proportion of career protection in the form of contract tenure (\emph{estatut'{a}rio}). Moving down the hierarchy, it is clear that party members predominantly occupy high level positions, with a much larger prevalence of party members in bureaucrat low, frontline high and then frontline low, as compared to other positions in the bureaucracy.

\begin{figure}[H]
    \centering
    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_cbo_group.pdf}
    \includegraphics[width = 6cm, height = 6cm]{figures/partisanship/plot_partisanship_by_contract_type.pdf}
    \caption{\textbf{Proportion of party-affiliated members}. On the left, breakdown of party membership by types of occupation. On the right, by type of contract.}
    \label{fig:partisan_breakdown}
\end{figure}

Additionally, party members are predominantly concentrated in career positions, which are protected from dismissals and departures. In contrast, temporary and other regular types of contract subject to turnover are relatively more protected from partisan influence. This empirical finding has substantive implications with respect to the nature of patronage in Brazil and developing countries more broadly. A large body of literature around civil service systems has proposed that establishing a career track with autonomy from political processes is a key development in protecting bureaucracies from political influence \citep{grindle2012jobs,carpenter2020forging}.

However, just because there are well established forms of career service does not mean that partisan affiliation is necessarily filtered out in the selection process. In particular, career services, with the exception of a formal requirement to do some form of public examination, does not necessarily prevent bureaucrats from joining a political party - and seldom do so. In fact, career services only structure the progression or retention of bureaucrats once they are in civil service, meaning that even if these positions are in theory autonomous from politically motivated turnover, these mechanisms do not prevent the selection of individuals who are already affiliated to a party from joining the ranks of career bureaucrats.
